Equalizer 2.0 --\\Convergence of a Parallel Rendering Framework

### 2018-02-08 paperIEEE.tex
Source footnote
git-svn-id: https://dev.ifi.uzh.ch/vmml/svn/Equalizer16@104 db7c848b-7755-4df2-b763-8796797afa52
Tweaks
git-svn-id: https://dev.ifi.uzh.ch/vmml/svn/Equalizer16@105 db7c848b-7755-4df2-b763-8796797afa52

\usepackage{mathptmx}
\usepackage{graphicx}
\usepackage{times}
%\usepackage{tabularx}
\usepackage{amsmath}
%\usepackage{url}
\usepackage{caption,subcaption}
\usepackage{multirow}
\usepackage{paralist}
\usepackage{listings}

@@

We conducted performance benchmarks for the decomposition modes on a cluster using hexacore Intel Xeon E5-2620v3 CPUs (2.4GHz), 16 GB main memory, 4 GBit/s Ethernet and QDR Infiniband. GCC 4.8 has been used with CMake 3.7 release mode settings to compile the software stack.

We tested the decomposition modes with both volume (\fig{fBenchmarkEVolve}) and polygonal data  (\fig{fBenchmarkEqPly}) and used test scenes that allowed us to finely adjust the rendering load the system has to cope with. In both cases the scene comprised of two rows of instantiated, identical models (the 1mm David model when rendering polygonal data, the $1024^3$ Flower model when rendering volumes), with 30 models in total. Rendering was performed
% with the \textsf{Equalizer} example applications eqPly (for meshes) and eVolve (for volumes)
at 2560x1440. The camera is initially placed in the center of the scene, between the two rows (\fig{fBenchmarkEVolve} and \fig{fBenchmarkEqPly}), rendering only half of the model instances. It is then moved backward over the duration of 800 frames, steadily increasing the rendering load by revealing more models, until all 30 instances are visible.

@@

In the case of volume rendering we also measured the performance of decomposition modes under heterogeneous load, which was easily achievable by varying the number of volume samples used for each fragment (1-7), while rendering. This allowed for a consistent linear scaling of rendering load, which was randomly varied (\fig{fEqualizers}) either per frame or per node. Such a linear scaling of load per node corresponds to a scaling of resources, e.g., doubling the rendering load on a specific node corresponds to halving its available rendering resources. To the system this node would then contribute the value 0.5 in terms of {\em normalized compute resources}, as illustrated by \fig{fEqualizers} (left).

This figure gives an impression of how individual modes perform on heterogeneous systems. In this case the tree equalizer performs best (\fig{fEqualizers} (left)), as it allows us to a priori define how much {\em usage} it should make of individual nodes, i.e., bias the allocation of rendering time, in accordance with the expected load / the available compute resources. \fig{fEqualizers} (right), on the other hand, illustrates how the tested decomposition modes perform on a system where compute resources fluctuate randomly every frame, as can arguably be the case for, e.g., cloud-based applications. \FIXME{Should we only mention this in discussion?} For this scenario tile equalizer seems best suited, as it performs load balancing implicitly and does not assume coherence of available resources between frames.

\begin{figure}[ht]\center
  \includegraphics[width=.9\columnwidth]{images/benchmark_eVolve}
  \caption{\label{fBenchmarkEVolve} Benchmarking decomposition modes with Flower volume model at $1024^3$ unsigned byte (1GB).}
\end{figure}

\begin{figure}[ht]\center
  \includegraphics[width=.9\columnwidth]{images/benchmark_eqPly2}
  \caption{\label{fBenchmarkEqPly} Benchmarking decomposition modes with David statue at 1mm resolution (988MB).}
\end{figure}

@@

To benchmark the data distribution we used two datasets: The David statue at 2mm
resolution (\fig{fDavid2mm}) and 3D volumes of spike frequencies of an
electrical simulation of three million neurons (\fig{fVolumes}).

\begin{figure}[ht]\center
  \includegraphics[width=.49\columnwidth]{images/david2mm}
  \caption{\label{fDavid2mm}David statue at 2mm resolution (145MB)}
\end{figure}

@@

In this section we analyse how data distribution and synchronization performs in
real-world applications. We extracted the existing data distribution code from a
mesh renderer (eqPly) and a volume renderer (Livre) into a benchmark application
to measure the time to initially map all the objects on the render client nodes,
and to perform a commit-sync of the full data set after mapping has been
established. All figures observe a noticable measurement jitter due to other
services running on the cluster. The details of the benchmark algorithm can be
found in the implementation~\cite{eqObjectBench}. used the same data sets as in
the previous section. We ran the benchmark on up to eight physical nodes, that
is, after eight processes nodes start to run two processes per node, which share
CPU, memory and network interface bandwidth.

@@

\begin{IEEEbiography}[{
\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{images/stefan}}]{
Stefan Eilemann} Stefan is the technical manager of the Visualization Team in
the Blue
Brain Project. His work involves working towards large-scale visualization for
Exascale simulations; interactive integration of simulation, analysis and
visualization as well as flexible frameworks for data sharing and dynamic
allocation of heterogenous resources. He is the CEO and founder of Eyescale and
also the lead developer of the \textsf{Equalizer} parallel rendering framework.
He has worked in SGI's Advanced Graphics Division as the technical lead of
OpenGL Multipipe SDK - a toolkit for scalable, parallel visualization software
on shared memory supercomputers. He received his masters degree in Computer
Science from EPFL in 2015, and his Engineering Diploma in Computer Science in
1998. He is currently working towards a PhD in Computer Science at the
Visualization and MultiMedia Lab at the University of Zurich.
\end{IEEEbiography}

@@

\begin{figure*}[p!]\center
  \includegraphics[height=.19\textheight]{images/daviddist}
  \caption{\label{fDavidDist}Object mapping (left) and data synchronization
    time (middle, detail view right) for the David 2mm data set in
    \fig{fDavid2mm}}
\end{figure*}
\begin{figure*}[p!]\center
  \includegraphics[height=.19\textheight]{images/spikesdist}
  \caption{\label{fSpikesDist}Object mapping (left) and data synchronization
    time (middle, detail view right) for the Spike data set in
    \fig{fVolumes}, left}
\end{figure*}
\begin{figure*}[p!]\center
  \includegraphics[height=.19\textheight]{images/beechnutdist}
  \caption{\label{fBeechnutDist}Object mapping (left) and data synchronization
    time (middle, detail view right) for the beechnut data set in
    \fig{fVolumes}, right}
\end{figure*}
\begin{figure*}[p!]\center
  \includegraphics[height=.19\textheight]{images/ifdist}
  \caption{\label{fIFDist}Object synchronization using different network
    transports}
\end{figure*}

>>>>

\usepackage{mathptmx}
\usepackage{graphicx}
\usepackage{times}
%\usepackage{tabularx}
\usepackage{amsmath}
\usepackage{hyperref}
\usepackage{caption,subcaption}
\usepackage{multirow}
\usepackage{paralist}
\usepackage{listings}

@@

We conducted performance benchmarks for the decomposition modes on a cluster
using hexacore Intel Xeon E5-2620v3 CPUs (2.4GHz), \FIXME{GPUs}, 16 GB main
memory, 4 GBit/s Ethernet and QDR Infiniband. GCC 4.8 has been used with CMake
3.7 release mode settings to compile the software stack.

We tested the decomposition modes with both volume data (\fig{fBenchmarkEVolve})
and polygonal data  (\fig{fBenchmarkEqPly}) and used test scenes that allowed us
to finely adjust the rendering load the system has to cope with. In both cases
the scene comprised of two rows of instantiated, identical models with 30 models
in total. Rendering was performed at a resolution of 2560x1440. The camera is
initially placed in the center of the scene, between the two rows
(\fig{fBenchmarkEVolve} and \fig{fBenchmarkEqPly}), rendering only half of the
model instances. It is then moved backward over the duration of 800 frames,
steadily increasing the rendering load by revealing more models, until all 30
instances are visible.

@@

For volume rendering we also measured the performance of decomposition modes
under heterogeneous load, which was easily achievable by varying the number of
volume samples used for each fragment (1-7) while rendering. This allowed for a
consistent linear scaling of rendering load, which was randomly varied
(\fig{fEqualizers}) either per frame or per node. Such a linear scaling of load
per node corresponds to a scaling of resources, e.g., doubling the rendering
load on a specific node corresponds to halving its available rendering
resources. To the system this node would then contribute the value 0.5 in terms
of {\em normalized compute resources}, as illustrated by \fig{fEqualizers}
(left).

This figure gives an impression of how individual modes perform on heterogeneous
systems. In this case the tree equalizer performs best (\fig{fEqualizers}
(left)), as it allows us to a priori define how much {\em usage} it should make
of individual nodes, i.e., bias the allocation of rendering time in accordance
with the (simulated) compute resources. \fig{fEqualizers} (right), on the other
hand, illustrates how the tested decomposition modes perform on a system where
compute resources fluctuate randomly every frame, as can arguably be the case
for, e.g., cloud-based applications. For this scenario tile equalizer seems best
suited, as it performs load balancing implicitly and does not assume coherence
of available resources between frames.

\begin{figure}[ht]\center
  \includegraphics[width=.9\columnwidth]{images/benchmark_eVolve}
  \caption{\label{fBenchmarkEVolve} Flower volume model at $1024^3$ unsigned byte (1GB).}
\end{figure}

\begin{figure}[ht]\center
  \includegraphics[width=.9\columnwidth]{images/benchmark_eqPly2}
  \caption{\label{fBenchmarkEqPly} David statue at 1mm resolution (988MB).}
\end{figure}

@@

To benchmark the data distribution we used two datasets: The David statue at 2mm
resolution (as in \fig{fBenchmarkEqPly}, but in 2mm resolution) and 3D volumes
of spike frequencies of an electrical simulation of three million neurons
(\fig{fVolumes}).

@@

In this section we analyse how data distribution and synchronization performs in
real-world applications. We extracted the existing data distribution code from a
mesh renderer (eqPly) and a volume renderer (Livre) into a benchmark application
to measure the time to initially map all the objects on the render client nodes,
and to perform a commit-sync of the full data set after mapping has been
established. All figures observe a noticable measurement jitter due to other
services running on the cluster. The details of the benchmark algorithm can be
found in the
\htmladdnormallinkfoot{implementation}{https://github.com/Eyescale/Equalizer/tree/paper2018/tools/eqObjectBench}.
used the same data sets as in the previous section. We ran the benchmark on up
to eight physical nodes, that is, after eight processes nodes start to run two
processes per node, which share CPU, memory and network interface bandwidth.

@@

\begin{IEEEbiography}[{
\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{images/stefan}}]{
Stefan Eilemann} Stefan was the technical manager of the Visualization Team in
the Blue Brain Project. His work involves working towards large-scale
visualization for Exascale simulations; interactive integration of simulation,
analysis and visualization as well as flexible frameworks for data sharing and
dynamic allocation of heterogenous resources. He is the CEO and founder of
Eyescale and also the lead developer of the \textsf{Equalizer} parallel
rendering framework. He received his masters degree in Computer Science from
EPFL in 2015, and his Engineering Diploma in Computer Science in 1998. He is
currently working towards a PhD in Computer Science at the Visualization and
MultiMedia Lab at the University of Zurich. \end{IEEEbiography}

@@

\begin{figure*}[p!]\center
  \includegraphics[height=.19\textheight]{images/daviddist}
  \caption{\label{fDavidDist}Object mapping (left) and data synchronization
    time (middle, detail view right) for the David 2mm data set}
\end{figure*}
\begin{figure*}[p!]\center
  \includegraphics[height=.19\textheight]{images/spikesdist}
  \caption{\label{fSpikesDist}Object mapping (left) and data synchronization
    time (middle, detail view right) for the Spike data set in
    \fig{fVolumes}, left}
\end{figure*}
\begin{figure*}[p!]\center
  \includegraphics[height=.19\textheight]{images/beechnutdist}
  \caption{\label{fBeechnutDist}Object mapping (left) and data synchronization
    time (middle, detail view right) for the beechnut data set in
    \fig{fVolumes}, right}
\end{figure*}
\begin{figure*}[p!]\center
  \includegraphics[height=.19\textheight]{images/ifdist}
  \caption{\label{fIFDist}Object synchronization using different network
    transports}
\end{figure*}
